---
title: "Netlogo vs. Julia programming: The best way to simulate opinion formation"
author-inputfile: "authors.tex" 
keywords: "Simulation \\and Agent-based modells \\and Programming languages \\and Netlogo \\and Julia programming."
bibliography: ["rpackages.bib", "bibliography.bib"]
biblio-style: "apalike"
link-citations: true
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
    template: lncs-template.tex
    md_extensions: +footnotes
    citation_package: biblatex
    dev: pdf
    
---


```{r knitr_init, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, paged.print=TRUE}
## Global options
library(knitr)
#options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               out.width='100%'
               )
opts_knit$set(width=100)

# Load required libraries
library(rmdformats)
library(tidyverse)
library(cowplot)
# add libraries here

# Add libraries to cite here:
pkgs <- c("knitr", "tidyverse", "rmdformats", "kableExtra", "scales", "psych", "rmdtemplates")
rmdtemplates::write_bib(pkgs)
```


```{r r_citations, eval=TRUE, include=TRUE, results="asis"}
rmdtemplates::line_cite(pkgs) # This creates a single line citing all packages
#rmdtemplates::list_cite(pkgs) # This creates a "thightlist" of all packages 
```

# Abstract
Hallo

<!--- Ca. 1/2 bis 1 Seite --->
# Introduction
When humans interact with each other or with digitized technology we speak of complex systems. The interaction of humans in such systems, for example in opinion-forming processes, leads to consequences that we cannot yet overlook or understand. An important component of socio-technical complex systems are single individuals that appear as human-in-the-loop [@Valdez2018human]. To look at the humans, their interactions and the resulting overall behavior, different simulation approaches using different programming languages are chosen.  

<!--- Since the spread of the Internet, different programming languages have also evolved.
Our world is more complex than years ago. Understanding what leads to complexity requires to consider the underlying systemic structure (Calero Valdez and Ziefle, 2018).
--->
<!--- Ca. 3 bis 4 Seiten --->
# Related Work
In this study, using Agent-based modelling we consider opinion formation processes, thus we look at a complex system. We want to know, whether it is possible tp create an Agent-based model with the programming language Netlogo and the programming language Julia. We further consider, how the two languages differ, which are the strenghths for creating Agent-based models of each programming language and which are the X. Contentwise, we built a bounded rationality model to simulate opinion formation. 

Therefore, we explain, which aspects lead to complexity, we introduce the method Agent-based modelling and the two programming languages Netlogo and Julia. Besides, we eplain what is knwon in theory about opinion formation or the spread of information. 

## Complexity
When examining opinion-forming processes, we look at a complex system. Such complex systems can be divided into several ontological levels or interacting subsystems on a micro- or macro-level [@Conte2012]. Further, complex systems lead to emergent phenomena. These complex systems and emergent phenomena are difficult to understand, because while it is easy to observe the individual system components, the resulting overall system cannot be considered as the sum of its parts. Instead, understanding the system behavior requires more than understanding the individual parts of the system [@Valdez2018human].  

To analyse complex systems we need a suitable approach, such as simulations, which enable to model the individual parts of a system and thus make the overall behavior visible. For the simulation of complex systems, Agent-based models are very well suited [@Epstein2007]. 


## Complex systems/Complexity
## Agent-based modelling
Agent-based models always consist of the agents or individuals and the environment in which the agents reside [@Bonabeau2002Agentbased]. They are neither a representation of reality, nor fully realistic or even complete.  Instead, they show a simplified reality. Nevertheless, agent-based models show behaviour on an individual level close to reality. By mapping the individual behavior, the behavior of the overall system can then be qualitatively observed [@Rouchier2008]. Agent-based models are well suited to replicate data and present the results to non-experts [@Kiesling2012]. The use of a method always requires an evaluation of the method. Evaluating agent-based models is not easy. In order to evaluate them, independent replicating and comparing with other model as well as a validation are necessary [@Rouchier2008].

The basis of agent-based modeling is the single agent or the individual. This agent is modelled programmatically as a template. In simulation, due agents make their own decisions based on how they perceive the environment in which they are situated. Thus, in our model
The perceptions of an agent usually determine the behavioral intent of the agent. If the agents are in a social network, as in our model, they influence their neighbours in the next iteration by their behavioural intention or the behaviour they show. To determine the probability of organizational acceptance, we analyze the results of several agent-based simulations.

A simple way to create agent-based models is to use software toolkits developed for the creation of simulations. These include the Netlogo toolkit considered in this study. With the use of such toolkits, it is easy to formulate the behavior of the individual agents. They also usually contain some useful interfaces. The interfaces allow to visualize the simulation states, interact with the simulation parameters and export the simulation results. In addition, they usually contain a batch mode. This is used to run a large number of simulations. Optimization strategies, such as genetic algorithms, help to find the most suitable parameters [@Calero2019].

 
To create agent-based models, Netlogo [@Wilensky1999] is the language most commonly used. Nevertheless, there are some other programming languages that are also suitable for creating agent-based models and that seem to be partly more intuitive, at least for people with programming experience. Therefore, in this study we compare two programming languages with respect to their suitability for creating agent-based models. 
## Netlogo
## Julia
## Opinion formation/spread of information


<!--- Ca. 3 bis 4 Seiten --->
# Method
Using two different programming languages (Julia language and Netlogo), we created two identical agent-based models that simulate opinion formation. Since our primary aim was to find out whether agent-based models could be implemented equally well in the two programming languages, we chose the most basic model of opinion-forming: bounded rationality. 

In our bounded rationality model, we defined the maximum number of agents, the maximum steps of the simulation, the seed and an epsilon in the beginning. The epsilon indicates how different the opinions of two people can be, so that they still include the other person's opinion in their opinion making. We further defined from the beginning, that each agent has an (floating) opinion between 0 and 1. In each simulation step, every agent compares his opinion with the opinion of an other agent. For example, if Anna compares her opinion with Ralf and the distance between the opinion of Anna and Ralf is smaller than the defined epsilon, then the two converge in their opinions.

We build the agent based models using either the Julia programming language or Netlogo. For the following analysis of the results we used R Markdown. 

## Netlogo
In this study, we compared two programming languages. We used version 6.0.4 of the multi-agent programming language Netlogo, which was developed by Uri Wilensky [@Wilensky1999]. Netlogo is visually appealing because the user is provided with a user interface to create and test an agent-based model. In addition to the user interface that simplifies the initial creation of a model, Netlogo also includes a batch mode that allows the user to run any number (usually several hundred) of simulations. 

## Julia

## Bounded rationality model
Since our primary goal was to compare the two programming languages with each other, we designed the parameters of the Netlogo model and the Julia model the identical way. Thus, we increased the comparability of the results of both models and reduced the complexity as much as possible. Table Y shows the settings, that we determined from the beginning of the model.   

### Opinion

### Network

## Before the model starts

## What do we compare
To find out whether both programming languages are euqually suitable to simulate our bounded rationality model, we look at severeal measurable criteria. These criteria include the outcomes and performance of both models. They further include how many lines of code are necessary to program the simulation. Another aspect, that we take into consideration, is, if learning Julia and Netlogo is equally difficult. For this aspect we consider both computer scientists who are familiar with other programming languages and a person who has no previous experience with programming languages. We further compare the explorability and scalability of both languages.

<!--- Ca. 4 bis 5 Seiten --->
# Results
Following, we first present the outcome of the bounded rationality model created with julia. Afterwards we show, if the model created with Netlogo showed the same or different results. Based on these results, we compare the two considered programming languages and show their advantages and disadvantages.

```{r results,child='analysis.Rmd', eval=FALSE}

```


## Julia simulation

## Comparison of Julia and Netlogo
When comparing both programming languages to create an agent-based model that simulates the bounded rationality model, Julia  proved to be a faster language less code lines were required to simulare our model. In addition, Julia turned out to be easier to learn for people with previous programming skills. In contrast, Netlogo proved to be a language that is easier to learn for people without programming skills.

<!--- Ca. 1 Seite --->
# Discussion
In our study, no language turned out to be the perfect programming language for creating agent-based models, but the choice of language seems to be a trade-off between various advantages and disadvantages. Of course, we have only focused on one very simple bounded rationality model, so that we would have to create further simulations with both languages to be able to make statements about the generality. 

<!--- Ca. 1/2 bis 1 Seite --->
# Conclusion and outlook
The results of our research have shown that Julia is the better choice for creating an agent-based model in some aspects, but also that Netlogo is better suited in some aspects. In the future, we plan to create more agent-based models using both programming languages, thus extending our comparison. 

# References


